\documentclass{article}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{braket, physics}
\usepackage{hyperref, cleveref}
\usepackage{algorithm}
\usepackage{xcolor}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\ga}[1]{{\noindent \textcolor{purple}{\emph{(GA: #1)}}}{}}

\title{Homework 4 Solution Key}
\author{Yi Lee}

\DeclareMathOperator{\XOR}{XOR}

\begin{document}

\include{macros}

\maketitle

\section*{Problem 1}

We will show that the construction is not secure if, roughly speaking, $G$ ignores the final bit of its input.
Formally, given any PRG $H:\set{0, 1}^n\rightarrow\set{0, 1}^\ell$,
consider the function $H':\set{0, 1}^{n+1}\rightarrow\set{0, 1}^\ell$ defined to: \begin{enumerate}
    \item On input $x\in\set{0, 1}^{n+1}$, split it as $x=a || b$ where $a$ contains the first $n$ bits of $x$.
    \item Output $H(a)$.
\end{enumerate}
We will show that $H'$ is a PRG, and that the mac $\mathsf{MAC}_k(m)=H'(k || m)$ is not one-time secure.
First, to show that $H'$ is a PRG, notice that $H'(s)$ and $H(s)$ are identically distributed, so for any distinguisher $\mathcal{D}$ we have
$\Pr[\mathcal{D}(H'(s))=1]=\Pr[\mathcal{D}(H(s))=1]$.
It therefore follows that
\begin{align*}
    \abs{\Pr[\mathcal{D}(H'(s))=1] - \Pr[\mathcal{D}(r)=1]} =
    \abs{\Pr[\mathcal{D}(H(s))=1] - \Pr[\mathcal{D}(r)=1]}
\end{align*}
is negligible for any distinguisher $\mathcal{D}$ due to the security of $H$.
Now, to show that the MAC is not one-time secure, consider the following adversary:
\begin{enumerate}
    \item Send the message $m=0^n1$ to the challenger and receive the reply $t$.
    \item Output the forgery $(0^{n+1}, t)$.
\end{enumerate}
By construction, the forgery succeeds with certainty.

\section*{Problem 2}

Suppose for the sake of contradiction that the MAC is \emph{not} one-time secure.
In other words, there exists an adversary $\mathcal{A}$ and a polynomial $q(n)$ such that
$\mathsf{OTForge}(\Pi, \mathcal{A})\geq\frac{1}{p} + \frac{1}{q(n)}$,
where $\Pi$ is the MAC scheme provided.

Consider then the following reduction $\mathcal{R}^\mathcal{A}(x)$ to the PRG game:
\begin{enumerate}
    \item Unpack the input $(a, b)\gets x$.
    \item Receive $m$ from $\mathcal{A}$.
    \item Evaluate $t\gets(a\cdot m+b \mod p)$ and send it to $\mathcal{A}$.
    \item Receive the forgery $(m^*, t^*)$ from $\mathcal{A}$.
    \item Evaluate $t'\gets(a\cdot m^* + b \mod p)$.
    \item If $m\ne m^*$ and $t' = t^*$, return $1$. Otherwise, return $0$.
\end{enumerate}
First, as the experiment $\mathcal{R}^{\mathcal{A}}(G(s))$ is equivalent to $\mathsf{OTForge}(\Pi, A)$, we have
\begin{align*}
\Pr[\mathcal{R}^{\mathcal{A}}(G(s))=1]&= \Pr[\mathsf{OTForge}(\Pi, \mathcal{A})=1] \\
&\geq\frac{1}{p} + \frac{1}{q(n)}
\end{align*}
by assumption.
On the other hand, the experiment $\mathcal{R}^{\mathcal{A}}(r)$ is equivalent to $\mathsf{OTForge}(\Pi', \mathcal{A})$,
where $\Pi'$ is the one-time MAC scheme constructed and proven secure in the lecture.
We therefore have
\begin{align*}
\Pr[\mathcal{R}^{\mathcal{A}}(r)=1]&= \Pr[\mathsf{OTForge}(\Pi', \mathcal{A})=1] \\
&=\frac{1}{p}.
\end{align*}
Finally, putting everything together, we conclude that
\[
\abs{\Pr[\mathcal{R}^{\mathcal{A}}(G(s))=1] - \Pr[\mathcal{R}^{\mathcal{A}}(r)=1]}\geq \frac{1}{q(n)}.\]

\section*{Problem 3}

\subsection*{Part a}

Consider the following adversary.
\begin{enumerate}
    \item Send $(0^\ell, 1^\ell)$ to the oracle and receive the tag $t$.
    \item Set $m\gets (1^\ell, 0^\ell)$.
    \item Output the forgery $(m, t)$.
\end{enumerate}
By construction, this adversary wins with certainty. 

\subsection*{Part b}

Consider the following adversary.
\begin{enumerate}
    \item Send $(0^\ell, 1^\ell)$ to the oracle, receive the tag $t$.
    \item Unpack $(y_1, y_2)\gets t$
    \item Set $m=(1^\ell, 0^\ell)$ and $t'\gets (y_2, y_1)$.
    \item Output the forgery $(m, t')$.
\end{enumerate}
By construction, this adversary wins with certainty.

\subsection*{Part c}

Consider the following adversary.
\begin{enumerate}
    \item Send $(0^\ell, 0^{\ell-1}1)$ to the oracle and receive the tag $t$.
    \item Unpack $(y_1, y_2)\gets t$.
    \item Send $(0^{\ell - 1}1, 0^\ell)$ to the oracle and receive the tag $t_2$.
    \item Unpack $(z_1, z_2)\gets t_2$.
    \item Set the message as $m^*\gets 0^{2\ell}$ and forgery as $t^*\gets (y_1, z_2)$.
    \item Output the forgery $(m^*, t^*)$.
\end{enumerate}
By construction, this adversary wins with certainty.

\end{document}